Reproducible Research: Peer Assessment 1
========================================================

#####1. Read file & attach it

```{r, echo=TRUE}
data<-read.csv("./activity.csv", colClasses=c("numeric", "Date", "numeric"))
attach(data)
```

#####2. Make a histogram of total number of steps taken per day.

```{r fig.width=4, fig.height=4, echo=TRUE}
totalsteps <- aggregate(steps ~ date, data, sum)
hist(totalsteps$steps,breaks=20,main="Total number of steps taken per day",xlab="Number of steps")
```

#####3. Calculate & report the mean & median total number of steps taken per day.

```{r echo=TRUE}
meansteps<-mean(totalsteps$steps)
mediansteps<-median(totalsteps$steps)
meansteps
mediansteps
```
The mean total number of steps taken per day is `r meansteps`. The median number of steps taken per day is `r mediansteps`.

#####4. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r fig.width=4, fig.height=4, echo=TRUE}
intervalsteps <- aggregate(steps ~ interval, data, mean)
plot(intervalsteps$steps,type="l",ylab="Mean number of steps",xlab="5-minute interval",main="Mean steps taken in 5-minute intervals")
```

#####5. Determine which 5-minute interval on average across all the days in the dataset, contains the maximum number of steps

```{r fig.width=4, fig.height=4, echo=TRUE}
maxsteps<-max(intervalsteps$steps,na.rm=TRUE)
maxsteps
maxsubset<-subset(intervalsteps,steps==maxsteps)
maxsubset
```
The maximum mean number of steps taken was in interval number `r maxsubset` steps being the mean number of steps.

#####6. Calculate and report the total number of missing values in the dataset.

```{r echo=TRUE}
missingvalues<-which(is.na(data))
missingcount<-length(missingvalues)
missingcount
```
The number of missing values in the dataset is `r missingcount`.

#####7. Create a new data set with all of the missing values replaced with the mean for that interval.

```{r echo=TRUE}
library(plyr)
get.mean <- function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))
new.data<- plyr::ddply(data[1:3], .(interval), transform,
                          steps = get.mean(steps),
                          date = date,
                          interval = interval)
new.data <- new.data[order(new.data$date,new.data$interval),]
row.names(new.data) <- 1:nrow(new.data)
```


#####8. Make a histogram of the total number of steps taken each day & calculate and report the mean & median total number of steps taken per day.

```{r fig.width=4, fig.height=4, echo=TRUE}
new.totalsteps <- aggregate(steps ~ date, new.data, sum)
hist(new.totalsteps$steps,breaks=20,main="Total number of steps taken per day",xlab="Number of steps")

new.meansteps<-mean(new.totalsteps$steps)
new.mediansteps<-median(new.totalsteps$steps)
new.meansteps
new.mediansteps
```

After replacing NA values with the mean for that interval, the mean step count was `r new.meansteps` (as compared to `r meansteps`). The new median was `r new.mediansteps` (as compared to `r mediansteps`). This indicates that the substitution of NA values did not significantly change the mean and median calculations, which makes logical sense.
